---
title: "Midterm + More Practice!"
subtitle: "Lecture 10"
date: "2025-05-29"
format: 
  revealjs:
    output-file: 10-practice-slides.html
auto-stretch: false
---

```{r}
#| include: false
library(tidyverse)
```

## Announcements

-   I will not hold office hours on Tuesday, June 3rd: this is when you have your take home!
-   Replacement midterm review office hours: June 2nd (3:30 - 5:30)
-   Be in class and lab on Monday!!!!
-   Tomorrow: Review game!

## Midterm Exam 1 {.smaller}

Worth 20% of your final grade; consists of two parts:

-   **In-class**: worth 70% of the Midterm 1 grade;

-   **Take-home**: worth 30% of the Midterm 1 grade.

## Material {.smaller}

Everything we have done so far:

-   plotting data with `ggplot` and interpreting plots
-   computing and understanding summary statistics
-   transforming data (row/column/grouping operations)
-   pivoting & joining data
-   data types/classes
-   importing data
-   *Monday's class*: *data science ethics*

## In-class

::: incremental
-   All multiple choice

-   You get both sides of one 8.5" x 11" note sheet that you and only you created (written, typed, iPad, etc)
:::

. . .

::: callout-important
If you have testing accommodations, make sure I get proper documentation from SDAO and make appointments in the Testing Center ASAP. The appointment should overlap substantially with our class time if possible.
:::

## What should I put on my cheat sheet? {.smaller}

::: incremental
-   description of common functions;
-   examples of function usage;
-   description of different visualizations: how to interpret, and what to use when;
-   doodles;
-   cute words of affirmation.
:::

. . .

::: callout-warning
Don't waste space on the details of any specific applications or datasets we've seen (penguins, Bechdel, gerrymandering, midwest, etc).
Anything we want you to know about a particular application will be introduced from scratch within the exam.
:::

## Example in-class question

Which command can replace a pre-existing column in a data frame with a new and improved version of itself?

a.  `group_by`
b.  `summarize`
c.  `pivot_wider`
d.  `geom_replace`
e.  `mutate`

## Example in-class question {.smaller}

```{r}
#| echo: false

df <- tibble(
  x = c(1, 2, 3, 4, 5, 6),
  y = c("Marie", "Marie", "Katie", "Mary", "Mary", "Mary")
)
```

::: columns
::: {.column width="40%"}
```{r}
df
```
:::

::: {.column width="50%"}
```{r}
#| eval: false 

df |>
  group_by(y) |>
  summarize(xbar = mean(x))
```

How many rows will this output have?

a.  1
b.  2
c.  3
d.  6
e.  11
:::
:::

## Example in-class question {.smaller}

::: columns
::: {.column width="50%"}
Which box plot is visualizing the same data as the histogram?
:::

::: {.column width="50%"}
```{r}
#| echo: false
#| warning: false
set.seed(100)
x <- c(rnorm(150), rnorm(350, mean = 2, sd = 0.5))
df <- tibble(x = x)
ggplot(df, aes(x = x)) +
  geom_histogram() + 
  theme(axis.text = element_text(size = 36)) + 
  labs(x = NULL, y = NULL)
```
:::
:::

```{r}
#| echo: false
#| fig-asp: 0.33
#| warning: false
w <- rnorm(500)
y <- c(pmax(rnorm(350, mean = -1), -2), rnorm(150, mean = 2, sd = 0.5))
z <- x + 10
X <- as_tibble(cbind(sort(rep(1:4, 500)), c(w, z, x, y))) |>
  mutate(
    V1 = case_when(
      V1 == 1 ~ "a",
      V1 == 2 ~ "b",
      V1 == 3 ~ "c",
      V1 == 4 ~ "d",
    )
  )
ggplot(X, aes(x = V2)) + 
  geom_boxplot() + 
  facet_wrap(.~V1, scales = "free") + 
  theme(axis.text = element_text(size = 18), 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  labs(x = NULL, y = NULL)
```

## Example in-class question {.smaller}

What code could have been used to produce `df_result`? Select all that apply.

::: columns
::: {.column width="25%"}
`df_X`

| state | year |
|-------|------|
| LA    | 2025 |
| NC    | 2025 |
| LA    | 2024 |
:::

::: {.column width="25%"}
`df_Y`

| state | region |
|-------|--------|
| LA    | south  |
| NC    | south  |
| CA    | west   |
:::

::: {.column width="50%"}
`df_result`

| state | year | region |
|-------|------|--------|
| LA    | 2025 | south  |
| NC    | 2025 | south  |
| LA    | 2024 | south  |
:::
:::

<br>

::: columns
::: {.column width="50%"}
a. `left_join(df_X, df_Y)`  
b. `right_join(df_X, df_Y)`  
c. `full_join(df_X, df_Y)`  
:::

::: {.column width="50%"}
d. `anti_join(df_Y, df_X)`  
e. `right_join(df_Y, df_X)`  
:::
:::
## Take-home

::: incremental
-   It will be just like a lab, but shorter;
-   Completely open-resource, but citation policies apply;
-   Absolutely no collaboration of any kind;
-   Seek help by posting *privately* on Ed;
-   Submit your final PDF to Gradescope in the usual way.
:::

## Reminder: conduct policies

::: incremental
-   Uncited use of outside resources or inappropriate collaboration will result in a zero and be referred to the conduct office;
-   If a conduct violation of any kind is discovered, your final letter grade in the course will be permanently reduced (A- down to B+, B+ down to B, etc);
-   If folks share solutions, all students involved will be penalized equally, the sharer the same as the recipient.
:::

## Things you can do to study

::: incremental
-   **Practice problems**: released Thursday February 13;
-   **Attend class tomorrow**: review game
-   **Old labs**: correct parts where you lost points;
-   **Old AEs**: complete tasks we didn't get to and compare with key;
-   **Code along**: watch these videos specifically;
-   **Textbook**: odd-numbered exercises in the back of IMS Chs. 1, 4, 5, 6
:::

# Let's Practice!

