---
title: "Multiple Linear Regression"
subtitle: "Lecture 15"
date: "2025-06-05"
format: 
  revealjs:
    output-file: 15-mult-reg-slides.html
---

```{r}
#| include: false
library(tidyverse)
library(tidymodels)
library(fivethirtyeight)
library(datasauRus)
movie_scores <- fandango |>
  rename(
    critics = rottentomatoes, 
    audience = rottentomatoes_user
  )

# Treat 'am' as a factor (categorical predictor)
mtcars$am <- factor(mtcars$am, labels = c("Automatic", "Manual"))

todays_ae <- "ae-12-modeling-penguins"

```

```{r}
#| include: false
movie_scores <- movie_scores %>%
  mutate(imdb_cat = case_when(
    imdb < 6.3 ~ "Low",
    imdb < 7.4 ~ "Med",
    TRUE ~ "High"
  ))
# Fit model: critic rating as outcome, audience level as predictor
model <- lm(audience ~ imdb_cat, data = movie_scores)
summary(model)

# Plot
ggplot(movie_scores, aes(x = imdb_cat, y = audience)) +
  geom_boxplot(fill = "orchid") +
  geom_jitter(width = 0.1, alpha = 0.5) +
  stat_summary(fun = mean, geom = "point", color = "darkred", size = 3) +
  labs(title = "Critic Score by Audience Rating Level",
       x = "Audience Rating Category",
       y = "Metacritic Score") +
  theme_minimal()
```

## Announcements

-   Office hours today: 1-3pm; old chem 203

-   Math warning: logarithms

-   Let me know if your Milestone 2 is ready to grade!

## Simple Linear Regression

-   A single quantitative $x$ and $y$

-   Population Model:

    $$
    Y = \beta_0 + \beta_1 X + \epsilon
    $$

-   Regression Line:

    $$
    \hat{Y} = b_0 + b_1 X
    $$

# Regression with Categorical Variables

## Regression with Categorical Variables {.smaller}

*It's possible to want to model an outcome based on a categorical predictor:*

$$ Y = \beta_0 + \beta_1 X + \epsilon$$

What does this even mean when $x$ is categorical ???

# Movie Example

## Example Data

```{r}
#| echo: false
movie_scores |>
  select(critics, audience, imdb_cat)
```

## Example Data: Visualization {.smaller}

```{r}
#| echo: false

ggplot(movie_scores, aes(x = imdb_cat, y = audience)) +
  geom_boxplot(fill = "orchid") +
  #geom_jitter(width = 0.1, alpha = 0.5) +
  #stat_summary(fun = mean, geom = "point", color = "darkred", size = 3) +
  labs(title = "Critic Score by Audience Rating Level",
       x = "Audience Rating Category",
       y = "Metacritic Score") +
  theme_minimal()
```

## Let's just try it... {.smaller}

```{r}
movies_fit <- linear_reg() |>
  fit(audience ~ imdb_cat, data = movie_scores)

tidy(movies_fit)
```

What's going on??

## What's going on? {.smaller}

Variable `imdb_cat` can take values *high, med,* or *low.* We want to model variable `audience` based on `imdb_cat`.

. . .

We tell R to fit `audience ~ imdb_cat` ... now what?

. . .

[We get ***dummy variables!!!***]{style="background-color: #fcefb3"}

## Dummy Variables {.smaller}

::: columns
::: {.column width="50%"}
| audience | imdb_cat |
|----------|----------|
| 70       | high     |
| 40       | low      |
| 50       | med      |
| 80       | high     |
:::
:::

## Dummy Variables {.smaller}

::: columns
::: {.column width="50%"}
| audience | imdb_cat |
|----------|----------|
| 70       | high     |
| 40       | low      |
| 50       | med      |
| 80       | high     |
:::

::: {.column width="50%"}
| audience | imdb_cat | high | low | med |
|----------|----------|------|-----|-----|
| 70       | high     |      |     |     |
| 40       | low      |      |     |     |
| 50       | med      |      |     |     |
| 80       | high     |      |     |     |
:::
:::

. . .

::: callout-important
In a given row, only *one* of the dummy variables for a given categorical variable can equal 1.
:::

## Write the regression output: {.smaller}

Using the categorical variable?

$$
\widehat{audience} = b_0 + b_1 \times imdb\_cat
$$

<br><br><br>

❌ WRONG ❌ 🛑 DON'T DO THIS 🛑 ❗SERIOUSLY ❗💔 Don't make me sad :( 💔

## What will a regression output look like? {.smaller}

Use the dummy variables!!

$$
\widehat{audience} = b_0 + b_1 \times low + b_2  \times med
$$

## What will a regression output look like? {.smaller}

Use the dummy variables!!

$$
\widehat{audience} = b_0 + b_1 \times low + b_2 \times med
$$

<br><br><br>

::: callout-important
All levels except the **baseline level** will show up in the output.
Here, high is the **baseline level.** You will be able to identify the baseline by looking at the data and the output!
:::

## Interpret the output: {.smaller}

$$
\widehat{audience} = b_0 + b_1 \times low + b_2 \times med
$$

-   (Intercept) $b_0$ ?

<br>

-   (Slope) $b_1$ ?

<br>

-   (Slope) $b_2$ ?

## Let's plug in some numbers: {.smaller}

```{r}
movies_fit <- linear_reg() |>
  fit(audience ~ imdb_cat, data = movie_scores)

tidy(movies_fit)
```

<br>

$$
\widehat{audience} = 85.1 -45.8 \times low -21.6 \times med
$$

## Interpretation with numbers: {.smaller}

$$
\widehat{audience} = 85.1 -45.8 \times low -21.6 \times med
$$

<br>

-   Expected audience score of high IMDB rating movies: 85.1

-   Movies with low IMDB ratings are expected to have, on average, audience scores 45.8 points lower than movies with high IMDB ratings

-   Movies with medium IMDB ratings are expected to have, on average, audience scores 21.6 points lower than movies with high IMDB ratings

# In General

## Generally: {.smaller}

Predicting [continuous outcome]{style="background-color: #fcefb3"} $y$ using [categorical explanatory variable]{style="background-color: #fcefb3"} $x$ with levels $cat_0, \ldots cat_L$:

## Generally: {.smaller}

Predicting [continuous outcome]{style="background-color: #fcefb3"} $y$ using [categorical explanatory variable]{style="background-color: #fcefb3"} $x$ with levels $cat_0, \ldots cat_L$:

$$
\hat{y} = b_0 + b_1 \times cat_1 + b_2 \times cat_2 \ldots + b_L \times cat_L
$$

## Generally: {.smaller}

Predicting [continuous outcome]{style="background-color: #fcefb3"} $y$ using [categorical explanatory variable]{style="background-color: #fcefb3"} $x$ with levels $cat_0, \ldots cat_L$:

$$
\hat{y} = b_0 + b_1 \times cat_1 + b_2 \times cat_2 \ldots + b_L \times cat_L
$$

-   $b_0$ : On average, expected $y$ for an entry of **baseline level** $cat_0$

<br>

-   $b_1$: On average, we expect the $y$ for an observation of level $cat_1$ to be be $b_1$ larger (smaller if negative) than a baseline $cat_0$ observation <br>

<br>

-   and so on....

## Back to AE-12! {.smaller}
