---
title: "AE 11: Modelling fish"
subtitle: "Suggested answers"
format:
  html:
    self-contained: true
---

For this application exercise, we will work with data on fish.
The dataset we will use, called `fish`, is on two common fish species in fish market sales.

```{r}
#| label: load-packages-data
#| message: false
library(tidyverse)
library(tidymodels)

fish <- read_csv("data/fish.csv")
```

The data dictionary is below:

| **variable**      | **description**        |
|-------------------|------------------------|
| `species`         | Species name of fish   |
| `weight`          | Weight, in grams       |
| `length_vertical` | Vertical length, in cm |
| `length_diagonal` | Diagonal length, in cm |
| `length_cross`    | Cross length, in cm    |
| `height`          | Height, in cm          |
| `width`           | Diagonal width, in cm  |

# Visualizing the model

We're going to investigate the relationship between the weights and heights of fish.

-   Create an appropriate plot to investigate this relationship. Add appropriate labels to the plot.

```{r}
#| label: plot-hw

ggplot(fish, aes(x = height, y = weight)) +
  geom_point() +
  labs(
    title = "Weights vs. lengths of fish",
    x = "Head-to-tail lentgh (cm)",
    y = "Weight of fish (grams)"
  )
```

-   If you were to draw a a straight line to best represent the relationship between the heights and weights of fish, where would it go?
    Why?

-   Now, let R draw the line for you.

```{r}
#| label: plot-hw-lm
ggplot(fish, aes(x = height, y = weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Weights vs. lengths of fish",
    x = "Head-to-tail lentgh (cm)",
    y = "Weight of fish (grams)"
  )
```

-   What types of questions can this plot help answer?

    *Is there a relationship between fish heights and weights of fish?*

```{=html}
<!-- -->
```
-   We can use this line to make predictions. Predict what you think the weight of a fish would be with a height of 10 cm, 15 cm, and 20 cm. Which prediction is considered extrapolation?\
    \
    *At 10 cm, we estimate a weight of 375 grams. At 15 cm, we estimate a weight of 600 grams At 20 cm, we estimate a weight of 975 grams. 20 cm would be considered extrapolation.*

```{=html}
<!-- -->
```
-   What is a residual?\
    \
    *Difference between predicted and observed.*

# Model fitting

-   **Demo:** Fit a model to predict fish weights from their heights.

```{r}
#| label: fit-model
fish_hw_fit <- linear_reg() |>
  fit(weight ~ height, data = fish)

fish_hw_fit
```

-   Predict what the weight of a fish would be with a height of 10 cm, 15 cm, and 20 cm using this model.

```{r}
#| label: predict-by-hand
x <- c(10, 15, 20)
-288 + 60.92 * x
```

-   **Demo:** Calculate predicted weights for all fish in the data and visualize the residuals under this model.

```{r}
#| label: residuals
fish_hw_aug <- augment(fish_hw_fit, new_data = fish)

ggplot(fish_hw_aug, aes(x = height, y = weight)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +  
  geom_segment(aes(xend = height, yend = .pred), color = "gray") +  
  geom_point(aes(y = .pred), shape = "circle open") + 
  theme_minimal() +
  labs(
    title = "Weights vs. heights of fish",
    subtitle = "Residuals",
    x = "Height (cm)",
    y = "Weight (gr)"
  )
```

# Model summary

-   **Demo:** Display the model summary including estimates for the slope and intercept along with measurements of uncertainty around them. Show how you can extract these values from the model output.

```{r}
#| label: model-summary
fish_hw_tidy <- tidy(fish_hw_fit)
fish_hw_tidy
```

-   **Demo:** Write out your model using mathematical notation.

Add response here.

# Correlation

We can also assess correlation between two quantitative variables.

-   What is correlation? What are values correlation can take?

Add response here.

-   Are you good at guessing correlation?
    Give it a try!
    <https://www.rossmanchance.com/applets/2021/guesscorrelation/GuessCorrelation.html>

-   **Demo:** What is the correlation between heights and weights of fish?

```{r}
#| label: correlation
fish |>
  summarize(r = cor(height, weight))
```

# Adding a third variable

-   **Demo:** Does the relationship between heights and weights of fish change if we take into consideration species? Plot two separate straight lines for the Bream and Roach species.

```{r}
#| label: plot-hws-lm
ggplot(fish, 
       aes(x = height, y = weight, color = species)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Weights vs. heights of fish",
    x = "Height (cm)",
    y = "Weight (gr)"
  )
```

# Fitting other models

-   **Demo:** We can fit more models than just a straight line. Change the following code below to read `method = "loess"`. What is different from the plot created before?

```{r}
#| label: plot-hw-loess
ggplot(fish, 
       aes(x = height, y = weight)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(
    title = "Weights vs. heights of fish",
    x = "Height (cm)",
    y = "Weight (gr)"
  )
```
