---
title: "AE 12: Modelling penguins"
---

In this application exercise we will be studying penguins.
The data can be found in the **palmerpenguins** package and we will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

You've seen the `penguins` data far too much at this point (sorry!!!), but we're going to bring it back one more time.

```{r}
#| label: glimpse-penguins
glimpse(penguins)
```

# Part 1

Our goal is to understand better **how various body measurements and attributes of penguins relate to their body mass**.
First, we are going to investigate the relationship between a penguins' flipper lengths and their body masses.

-   *Based on our research focus, body mass is the response variable.*

**Task 1 - Exploratory Data Analysis:** Visualize the relationship between flipper length and body mass of penguins.

```{r}
#| label: plot-bm-fl

# Add code here!
```

## Correlation

**Task 2 - Complete the following**

-   **Question:** What is correlation? What values can correlation take?

    *Add answer here. \
    *

    ::: callout-note
    *Are you good at guessing correlation?* [Give it a try](https://www.guessthecorrelation.com) with this game!
    :::

-   **Code:** What is the correlation between flipper length and body mass of penguins?

```{r}
#| label: correlation

# add code here
```

## Defining, fitting, and summarizing a model

**Task 3 (Demo) :** Write the [*population model*]{.underline} (model for the *true values)* below that explains the relationship between body mass and flipper length.

*Add answer here!*

**Task 4:** Fit the linear regression model and display the results.
[*Write the estimated model output below.*]{.underline}

```{r}
#| label: fit-bm-fl

# add code here
```

*Writee model output here.*

**Task 5:** Interpret the slope and the intercept in the context of the data.

-   *Intercept*:

-   *Slope*:

**Task 6:** Recreate the visualization from above, this time adding a regression line to the visualization `geom_smooth(method = "lm")`.

```{r}
#| label: plot-bm-fl-line

# add code here
```

**Task 7 (Demo):** What is the estimated body mass for a penguin with a flipper length of 210?

```{r}
#| label: predict-210

# add code here
```

**Task 8:** What is the estimated body mass for a penguin with a flipper length of 100?
Add code to find it!
Is there anything weird about making this prediction?

```{r}
#| label: predict-100

# add code here
```

# Part 2: Another model

**Task 9:** A different researcher wants to look at body weight of penguins based on the island they were recorded on.

-   **Question:** How are the variables involved in this analysis different?\

    \
    *Add answer here.*

-   **Code:** Make an appropriate visualization to investigate this relationship below.
    Additionally, calculate the mean body mass by island.

```{r}
#| label: plot-bm-island

# add plot here
```

```{r}
#| label: mean-bm-island

# add mean by island
```

**Task 10:** Change the geom of your previous plot to `geom_point()`.
Use this plot to think about how R models these data.

```{r}
#| label: plot-bm-island-points

# add code here
```

**Task 11:** Fit the linear regression model and display the results.
Write the estimated model output below.

```{r}
#| label: fit-bm-island

# add code here
```

***What looks different about these results than what we saw before? How can we interpret them?***

::: callout-note
Let's go back to the slides now - we'll learn a little more before continuing.
:::

## Interpreting Categorical Predictors

**Task 12:** Fill in the blanks.

-   Intercept: Penguins from \_\_\_\_\_\_\_\_\_ island are expected to weigh, on average, \_\_\_\_\_\_\_ grams.

-   Slopes:

    -   Penguins from \_\_\_\_\_\_\_\_\_ are expected to weigh, on average, \_\_\_\_\_\_\_ grams \_\_\_\_\_ than those from \_\_\_\_\_\_\_.

    -   Penguins from \_\_\_\_\_\_\_\_\_ island are expected to weigh, on average, \_\_\_\_\_\_\_\_ grams \_\_\_\_\_\_\_ than those from \_\_\_\_\_\_\_.

**Task 13:** What is the estimated body weight of a penguin on Biscoe island?
What are the estimated body weights of penguins on Dream and Torgersen islands?

```{r}
#| label: predict-bm-island

# add code here
```
